Converging High-Throughput and High-Performance Computing: A Case Study

### 2017-08-28 Experiments.tex
Comments to address before submission

\ifreview
Comments to address in this section:
\begin{enumerate}
    \item REVIEWER 2: In the weak scalability evaluations (in Figure 7 and
    8), the overhead of the Next Generation Executor (NGE) has grown linearly
    with the number of the nodes. Can you explain why the average time
    execution time of AthenaMP is decreased even when the overhead is
    increasing?
    I checked the data. Figure 8 was simply wrong. I think that we have overwritten the original with an older version.  Figure 7: I used a smaller integration step to compute the integral of the pilot concurrency. It gets better but not as much as I would like. Our data has something sketchy.  
\end{enumerate}
\fi

\subsection{Experiments}
\label{sec:ngeExp}

@@

\subsubsection{Weak scalability}

In this experiment we run as many AthenaMP instances (hereafter referred to
as tasks) as the number of nodes controlled by the pilot. Each AthenaMP
simulates 100 events, requiring $\sim 4200$ seconds on average.

@@

\begin{figure}[!t]
        \includegraphics[height=4.5cm,width=\columnwidth]{./figures/NGE/weak1.pdf}
   	\vspace{-0.3in}
    \caption{Weak scalability: average pilot duration, average duration of one AthenaMP execution, and pilot's overhead as a function of pilot sizes (200, 500, 1000 and 2000 nodes).}
\label{fig:weakScal1a}
\end{figure}

@@

\subsubsection{Weak scalability with multiple generation}

@@

\begin{figure}[!t]
        \includegraphics[height=4.5cm,width=\columnwidth]{./figures/NGE/weak2.pdf}
        \vspace{-0.3in}
    \caption{Weak scalability with multiple generations (where each generation has approximately 1/6th the number of events compared to Fig. 7): average pilot
    duration, average duration of sequential AthenaMP executions, and
    pilot's overhead for pilot with 256, 512, 1024 and 2048 nodes.}
\label{fig:weakScal2a}
\end{figure}

@@

\subsubsection{Strong scalability}

@@

\begin{figure}[!t]
        \includegraphics[height=4.5cm,width=\columnwidth]{./figures/NGE/strong.pdf}
        \vspace{-0.3in}
    \caption{Strong scalability: Average pilot duration, average duration of
    sequential AthenaMP executions, and pilot's overhead for pilots with 256, 512, 1024 and 2048 nodes.}
\label{fig:strongScala}
\end{figure}

>>>>

\ifreview
Comments to address in this section:
\begin{enumerate}
    \color{red} 
    \item REVIEWER 3: The scalability benchmarks in section "V.B Experiments"
    should be better explained: what *extacly* is being kept constant in each
    experiment, and what is being varied and over what range. Also the
    captions in the figures do not help much.
\end{enumerate}
\fi

% ---------------------------------------------------------------------------
\subsection{Experiments}\label{sec:ngeExp}

@@

% ---------------------------------------------------------------------------
\subsubsection{Weak scalability}

In this experiment we run as many AthenaMP instances (hereafter referred to
as tasks) as the number of nodes controlled by the pilot. Each AthenaMP
simulates 100 events, requiring \(\sim 4200\) seconds on average.

@@

\begin{figure}[!t]
    \includegraphics[height=4.5cm,width=\columnwidth]{./figures/NGE/weak1.pdf}
   	\vspace{-0.3in}
    \caption{Weak scalability: average pilot duration, average duration of
    one AthenaMP execution, and pilot's overhead as a function of pilot sizes
    (200, 500, 1000 and 2000 nodes).}\label{fig:weakScal1a}
\end{figure}

@@

% ---------------------------------------------------------------------------
\subsubsection{Weak scalability with multiple generation}

@@

\begin{figure}[!t]
    \includegraphics[height=4.5cm,width=\columnwidth]{./figures/NGE/weak2.pdf}
    \vspace{-0.3in}
    \caption{Weak scalability with multiple generations (where each
    generation has approximately 1/6th the number of events compared to Fig.
    7): average pilot duration, average duration of sequential AthenaMP
    executions, and pilot's overhead for pilot with 256, 512, 1024 and 2048
    nodes.}\label{fig:weakScal2a}
\end{figure}

@@

% ---------------------------------------------------------------------------
\subsubsection{Strong scalability}

@@

\begin{figure}[!t]
    \includegraphics[height=4.5cm,width=\columnwidth]{./figures/NGE/strong.pdf}
    \vspace{-0.3in}
    \caption{Strong scalability: Average pilot duration, average duration of
    sequential AthenaMP executions, and pilot's overhead for pilots with 256,
    512, 1024 and 2048 nodes.}\label{fig:strongScala}
\end{figure}

### 2017-08-28 PandaEvolution.tex
Comments to address before submission

% \ifreview
% Comments to address in this section:
% \begin{enumerate}
% 	\item REVIEWER 2: The authors spent quite a few paragraphs for the detail
% 	design and implementation of the PanDA WMS, but introduced few background
% 	of the LHC applications. According to the paper, I understand the
% 	applications running on Titan are a large amount of small independent
% 	jobs. Thus, it seems less cost-effective to run these jobs on a
% 	world-leading supercomputer like Titan which has a fast and expensive
% 	interconnection. It will be more interesting if the authors can compare
% 	the solution on Titan (proposed in this paper) with the original
% 	distributed computing environment and HPECloud from Google mentioned in
% 	the related work section.
% 	\item REVIEWER 5: In Section 3.2, GPUs are not considered in this work,
% 	which would be a waste for using HPC computing. Could this work be
% 	extended to include GPUs or other heterogenous resources?
% \end{enumerate}
% \fi
The upcoming LHC Run 3 will require more resources than the Worldwide LHC
Computing Grid (WLCG) can provide. Currently, PanDA WMS uses more than
300,000 cores at more than 100 Grid sites, with a peak performance of 0.3
petaFLOPS\@. This capacity will be sufficient for the planned analysis and
data processing, but it will be insufficient for the Monte Carlo production
workflow and any extra activity. To alleviate these challenges, ATLAS is
expanding the current computing model to include additional resources such as
the opportunistic use of supercomputers.

@@

% ---------------------------------------------------------------------------
\subsection{Architectures, Interfaces and Workloads}
\label{ssec:panda-titan}

>>>>

\ifreview
Comments to address in this section:
\begin{enumerate}
	\color{red} 
	\item REVIEWER 2: the technical discussion of TITAN and WLCG differences
	is abruptly stopped by the description of Monte Carlo event generation
	and detector simulation (3A).
	\item REVIEWER 2: The single listing (3B) of similar projects within the
	collaboration does not detail whether this is derived or separate work.
	\item REVIEWER 2: How does the TITAN environment prevent usage of DTNs as
	thin proxies? Is connectivity possible from compute node to DTNs via MPI,
	message queues, sockets or other means? Are the restrictions a
	shortcomming of TITAN, or the majority of HPC environments? Given the
	proclaimed importance of HPC resources, a generic evaluation seems
	needed.
	\item REVIEWER 2: Why was the PanDA Broker designed orthogonal to the
	pilot model, performing both resource acquisition and workload execution?
	How do pilots conflict with backfill slots, when the former are designed
	to acquire resources independent of workload requirements? Why does each
	Broker instance query and use backfill separately? Given that this seems
	to be a major bottleneck, the choice of such a design begs for
	motivation.
	\item REVIEWER 2: Some facts about WLCG seem outdated:
	\begin{itemize}
		\item The number of 30,000 unused TITAN cores (§IIIB) is equated to
		roughly 10\% of the WLCG. However, WLCG currently has 591,282 cores
		[REBUSCPU]. https://wlcg-rebus.cern.ch/apps/capacities/federations/
		\item WLCG sites are claimed to provide only 8 cores per worker node
		(§IIIB). WLCG has a significant number of sites providing 16 or more
		cores per worker node [REBUSCPU]. https://wlcg-rebus.cern.ch/apps/capacities/federations/
	\end{itemize}
	\item REVIEWER 2: Schematic figures (Fig 1, Fig 2) are overly complex,
	while their captions are not sufficient to understand them. The numbering
	is only partially explained in the text.
\end{enumerate}
\fi

The upcoming LHC Run 3 will require more resources than the Worldwide LHC
Computing Grid (WLCG) can provide. Currently, PanDA WMS uses more than
300,000 cores at more than 100 Grid sites, with a peak performance of 0.3
petaFLOPS\@. This capacity will be sufficient for the planned analysis and
data processing, but it will be insufficient for the Monte Carlo production
workflow and any extra activity. To alleviate these challenges, ATLAS is
expanding the current computing model to include additional resources such as
the opportunistic use of supercomputers.

@@

% ---------------------------------------------------------------------------
\subsection{Architectures, Interfaces and Workloads}\label{ssec:panda-titan}

### 2017-08-28 conclusion.tex
Comments to address before submission

\ifreview
Comments to address in this section:
\begin{enumerate}
	\item REVIEWER 3: How extensible is this? Do you have a sense of backfill
	availability on other systems? Is this unique to Titan or Cray or more
	ubiquitous of an option?
\end{enumerate}
\fi

@@

>>>>

\ifreview
Comments to address in this section:
\begin{itemize}
	\color{red} 
	\item REVIEWER 2: How does Summit, the TITAN successor, compare to the
	conclusions? The end of the TITAN life-cycle basically invalidates the
	applicability of all conclusions if this generalisation is missing.
	\item REVIEWER 3: This statement in the "conclusions" section is quite
	exaggerated: "The experience captured in this paper will prove invaluable
	for designing WMS for computational campaigns and will provide a baseline
	to evaluate the relative merits of different approaches." WMS systems for
	LHC experiments have been developed and run in production for many years;
	while this paper adds interesting experience and data points, it's
	unlikely that this will be "invaluable" for new designs and a cornerstone
	for their evaluation.
\end{itemize}
\fi

@@

% The development of NGE is looking forward to deal with these challenges by
% providing an abstraction layer for HPC resources together with a set of
% APIs for the coordination and fast deployment of heterogeneous workloads.
% In this paper, we provided the results of a preliminary testing of the NGE.
% The testing pointed out the robustness of NGE at handling an high number of
% tasks and worker nodes by mantaining an overhead that grows linearly with
% the number of tasks.

### 2017-08-28 introduction.tex
Comments to address before submission

\ifreview
Comments to address in this section:
\begin{enumerate}
	\item REVIEWER 1: listing the key observations in one place, and in a
	consistent fashion. I found The 5 points listed in the first part of the
	rebuttal quite compelling, and strongly urge the authors to put them in
	the paper in a prominent location.
	\item REVIEWER 5: I would have preferred to see a more detailed list of
	contributions, including the novelty of the proposed approach as compared
	to prior work.
	\item I am ambivalent about the use of the term ``converging'' in the
	paper's title. While the paper does port workloads to high-performance
	computing resources, it does not deal with compute-oriented optimizations
	(parallelization), lowering communication, etc., which are often major
	concerns in HPC programming. In this sense, the paper is closer in spirit
	to how to port high-throughput apps to high-performance clusters. Agreed?
	{\bf  Need to clarify it is not the task that is being converged but the workload and execution model that is being converged}
	\item REVIEWER 1: A glossary would be welcomed. {\bf Not addressable}
\end{enumerate}
\fi

>>>>

\ifreview
Comments to address in this section:
\begin{enumerate}
	\color{red} 
	\item REVIEWER 1: The introduction is overly verbose and prosaic. The
	praise for LHC collaborations is off-putting, especially as it ignores
	other work. For example, claiming "CMS and ATLAS experiments utilize
	arguably the largest production grade distributed computing solutions"
	disregards the obvious commercial contenders. By blatantly excluding
	obvious comparisons, it raises the question how much other work is
	ignored.
\end{enumerate}
\fi

### 2017-08-28 nge.tex
Comments to address before submission

% As seen in \S\ref{ssec:panda_titan}, PanDA Broker is deployed on the DTNs
% as Titan's worker nodes lack connectivity to the wide area network.
The lack of pilot capabilities in PanDA Broker impacts both the efficiency
and the flexibility of PanDA's execution process. Pilots could improve
efficiency by increasing throughput and enabling greater backfill
utilization. Further, pilots would make it easier to support heterogeneous
workloads.

The absence of pilots 
% imposes the static coupling between MPI scripts submitted to the PBS batch
% system and detector simulations (\S\ref{sec:panda_titan}). This
makes the scheduling of multiple generations of workload on the same PBS job
impossible: once a statically defined number of detector simulations are
packaged into a PBS job and this job is queued on Titan, no further
simulations can be added to that job. New simulations have to be packaged
into a new PBS job that needs to be submitted to Titan based upon backfill
and PanDA Brokers availability.

>>>>

\ifreview
Comments to address in this section:
\begin{itemize}
	\color{red} 
	\item REVIEWER 3: The first paragraphs of section V. are basically a
	rediscovery of the advantages of pilot jobs for WLCG task scheduling.
	Since all those pain points were well-known in the WLCG community before
	the project started, it should be explained why a pilot job approach was
	not taken in the first place.
	\item REVIEWER 3: Figure 6. refers to a "User Workstation" as the running
	place for the Pilot and Unit Managers, which is never mentioned in the
	text. (I got the impression that the NGE would run on a DTN node; if this
	is not the case and separate workstations are being used for the testing
	phase, this should be made clear.)
\end{itemize}
\fi

The lack of pilot capabilities in PanDA Broker impacts both the efficiency
and the flexibility of PanDA's execution process. Pilots could improve
efficiency by increasing throughput and enabling greater backfill
utilization. Further, pilots would make it easier to support heterogeneous
workloads.

The absence of pilots makes the scheduling of multiple generations of
workload on the same PBS job impossible: once a statically defined number of
detector simulations are packaged into a PBS job and this job is queued on
Titan, no further simulations can be added to that job. New simulations have
to be packaged into a new PBS job that needs to be submitted to Titan based
upon backfill and PanDA Brokers availability.

### 2017-08-28 overall.tex
Checkpoint
Comments to address before submission

\ifreview
Comments to address in this section:
\begin{enumerate}
    % \item REVIEWER 2: In Figure 4, there are some inconsistencies. The
    % figure caption shows there are only 62175 measures, but in the scatter
    % plot (right) there are 62555 data points. In addition, according to the
    % distribution of the number of work nodes (left), it does not look like
    % the mean number of work available is 691 as the figure caption
    % illustrates.
    \item 
    % REVIEWER 3: It’s interesting that Feb/March have such an increase in
    % backfill availability in 2017. The authors attribute this to hardware
    % upgrades for the DTNs. It’s mentioned that the load is only 3.6\% now
    % and therefore further upgrades will likely not cause similar
    % improvement. What was the previous CPU load? 
    Why were particularly low percentages utilized in April/May 2016?
    % \item REVIEWER 4: “Genat4” at the end of Section 4 is a typo.
    % \item REVIEWER 4: Please explain why you used Titan at all, in spite of
    % your statement that it is 1/2 to 1/3 as efficient as other resources.
    % Was it merely that you were given so much time on this resource that
    % the inefficiency doesn’t matter?
    \item REVIEWER 4: Why didn’t you benchmark the performance of your
    workload on other HPC resources that might have delivered better
    performance, prior to reengineering your entire workflow to deal with the
    challenges posed by Titan, a machine that doesn’t seem to justify the
    pain?
    \item REVIEWER 4: The INCITE program operational requirements for the
    LCFs are 90\% uptime. Using large batches of Geant4 jobs that run on a
    single node or less to reach this metric sounds like gaming the system
    rather than an accomplishment.
    % \item REVIEWER 5: In Figure 3, it shows the unused core hours are much
    % higher than those used by ATLAS\@. What are the reasons for such low
    % usage?
\end{enumerate}
\fi

@@

On February 2017, PanDA Brokers used almost twice as much total backfill
availability than in any other month (preliminary results for March 2017
displayed in Fig.~\ref{fig:backfill-utilization} confirm this trend). No
relevant code update was made during that period and logs indicated that the
brokers were able to perform faster. This is likely due to hardware upgrades
on the DTNs. The absence of continuous monitoring of those nodes does not
allow to quantify bottlenecks but spot measurements of their load indicate
that a faster CPU and better networking were likely responsible for the
improved performance. Investigations showed an average CPU load of 3.6\% on
the upgraded DTNs, as opposed to the ``high'' utilization reported by OLCF
for the previous DTNs. As such, further hardware upgrades seem unlikely to
improve significantly the performance of PanDA Brokers.\mtnote{Asked Segey
and Danila for values previous HW upgrade}

>>>>

\ifreview
Comments to address in this section:
\begin{enumerate}
    \color{red} 
    \item REVIEWER 2: What are the reasons for the low efficiency of PanDA
    Broker? Why are only 30\% of backfill slots useable, if they are
    basically uncontested? Why does Fig 3 show an increase of absolute and
    relative efficiency in Aug-16, when the text claims this in Sep-16 due to
    more PanDA Brokers? Why are only 20 PanDA Brokers deployed if this is a
    limiting factor? Why is it questionable that more PanDA Brokers would
    increase efficiency? If only 5/6 of resources below the mean are
    unusable, this implies an efficiency of almost 60\% being reachable.
    \item REVIEWER 2: Given an expected increase in resource need of 10-100,
    how is the capacity of 4\% WLCG significant? How does this relate to the
    manpower required to develop and maintain a custom solution per HPC site?
    How many TITAN resources are actually available to HEP, given the use by
    other domains? The magnitude of capacities presented as available is in
    major contrast to the requirements sketched in the introduction.
    \item REVIEWER 2: How does the poor metadata read performance of
    1500-6300 s (25\%-100\% of minimum walltime!) affect efficiency? Is it
    factored into the choice of walltime requests? Is it applicable only to
    TITAN, to HPC in general, or even to regular WLCG sites? Are there
    solutions at application, e.g. the use of packaged python scripts and
    libraries? The information presented is insufficient to estimate whether
    this issue applies to other environments and usecases as well, or is
    simply due to suboptimal implementation/configuration.
    \item REVIEWER 3: The HEP-SPEC06 benchmark, which is used as a
    ``computing power'' measure in LCG is mostly integer-artihmetic based: FP
    operations accounts for \~10\% of the average workload (see [1], section
    ``Why do you use `all\_cpp' instead of SPECint or SPECfp?'') While this
    figure certainly comes from averaging over many different workloads
    across all experiments, it seems a bit surprising that FP scheduling
    contention is responsible for 50\% decrease in compute efficiency on
    Titan when compared to other WLCG sites (paper page 6).
\end{enumerate}
\fi

@@

On February 2017, PanDA Brokers used almost twice as much total backfill
availability than in any other month (preliminary results for March 2017
displayed in Fig.~\ref{fig:backfill-utilization} confirm this trend). No
relevant code update was made during that period and logs indicated that the
brokers were able to perform faster. This is likely due to hardware upgrades
on the DTNs. The absence of continuous monitoring of those nodes does not
allow to quantify bottlenecks but spot measurements of their load indicate
that a faster CPU and better networking were likely responsible for the
improved performance. Investigations showed an average CPU load of 3.6\% on
the upgraded DTNs, as opposed to the ``high'' utilization reported by OLCF
for the previous DTNs. As such, further hardware upgrades seem unlikely to
improve significantly the performance of PanDA Brokers.

### 2017-08-28 paper.tex
Comments to address before submission

% ---------------------------------------------------------------------------
% ABSTRACT
% ---------------------------------------------------------------------------
\maketitle
\begin{abstract}
\input{abstract}
\end{abstract}
% \keywords{Distributed computing, Workload management system}

% ---------------------------------------------------------------------------
% I - INTRODUCTION
% ---------------------------------------------------------------------------
\section{Introduction}
\label{sec:intro}
\input{introduction}

% ---------------------------------------------------------------------------
% II - PanDA OVERVIEW
% ---------------------------------------------------------------------------
\section{PanDA Overview}
\label{sec:panda_overview}
\input{PandaOverview}

% ---------------------------------------------------------------------------
% III - DEPLOYING PANDA ON A LEADERSHIP-SCALE SYSTEM
% ---------------------------------------------------------------------------
%\section{Deploying PanDA on a Leadership-scale system}
\section{Deploying PanDA on Titan}
\label{sec:panda_titan}
\input{PandaEvolution}

% ---------------------------------------------------------------------------
% IV - ANALYSIS AND DISCUSSION
% ---------------------------------------------------------------------------
\section{Analysis and Discussion}
\label{sec:panda_titan}
\input{overall}

% ---------------------------------------------------------------------------
% VI - THE NEXT GENERATION EXECUTOR
% ---------------------------------------------------------------------------
% \section{PANDA: Future/RoadMap}
\section{PANDA\@: The Next Generation Executor}
\label{sec:panda_roadmap}
\input{nge}
\input{Experiments}

% ---------------------------------------------------------------------------
% VII - RELATED WORK
% ---------------------------------------------------------------------------
%\vspace{-0.4in}
\section{Related Work}
\label{sec:related}
\input{bckground}

% ---------------------------------------------------------------------------
% VIII - CONCLUSION
% ---------------------------------------------------------------------------
%\vspace{-0.1in}
\section{Conclusion}
\label{sec:conclusion}
\input{conclusion}

% ---------------------------------------------------------------------------
% ACKNOWLEDGEMENTS
% ---------------------------------------------------------------------------
% \section*{Acknowledgements}
% \label{sec:ack}
%
% \input{ack}

@@

% ---------------------------------------------------------------------------
% A - APPENDIX
% ---------------------------------------------------------------------------
% \appendix
% \section{Artifact Description --- Converging High-Throughput and High-Performance Computing: A Case Study}
% \label{apnx:ac}
% \input{appendix}

>>>>

% ---------------------------------------------------------------------------
% ABSTRACT
% ---------------------------------------------------------------------------
\maketitle

\ifreview
General comments to address:
\begin{itemize}
    \color{red} 
    \item REVIEWER 1: There is a problem with capitalization in some
    reference titles, ie. [2], [3], [4], [7], [8], [9], [10], [11], [12],
    [13]. Reference [6] is just an author list, year and title. Should there
    be more here?
    \item REVIEWER 2:  Key concepts such as pilot jobs, tasks and workloads
    are not meaningfully defined.
    \item REVIEWER 2: the information provided is insufficient to asses in
    how far any of the presented evaluation, characterization or lessons are
    applicable outside of TITAN and PanDA Broker combined.
    \item REVIEWER 2: Several references are of poor quality. For example,
    [19] is a slide collection even though there are many papers for the
    referenced work.
    \item REVIEWER 3: Title looks too general for what amounts to the
    implementation of an adapter component in a system that is very specific
    to a single HEP experiment. (In general, title, abstract and conclusions
    seem to over-state the contributions of this paper, considering this is a
    general eScience conference and not a WLCG-specific one.)
\end{itemize}
\fi

\begin{abstract}
\input{abstract}
\end{abstract}
% \keywords{Distributed computing, Workload management system}

% ---------------------------------------------------------------------------
% I - INTRODUCTION
% ---------------------------------------------------------------------------
\section{Introduction}\label{sec:intro}
\input{introduction}

% ---------------------------------------------------------------------------
% II - PanDA OVERVIEW
% ---------------------------------------------------------------------------
\section{PanDA Overview}\label{sec:panda_overview}
\input{PandaOverview}

% ---------------------------------------------------------------------------
% III - DEPLOYING PANDA ON A LEADERSHIP-SCALE SYSTEM
% ---------------------------------------------------------------------------
\section{Deploying PanDA on Titan}\label{sec:panda_titan}
\input{PandaEvolution}

% ---------------------------------------------------------------------------
% IV - ANALYSIS AND DISCUSSION
% ---------------------------------------------------------------------------
\section{Analysis and Discussion}\label{sec:analysis}
\input{overall}

% ---------------------------------------------------------------------------
% V - THE NEXT GENERATION EXECUTOR
% ---------------------------------------------------------------------------
\section{PANDA\@: The Next Generation Executor}\label{sec:panda_roadmap}
\input{nge}
\input{Experiments}

% ---------------------------------------------------------------------------
% VI - RELATED WORK
% ---------------------------------------------------------------------------
\section{Related Work}\label{sec:related}
\input{bckground}

% ---------------------------------------------------------------------------
% VII - CONCLUSION
% ---------------------------------------------------------------------------
\section{Conclusion}\label{sec:conclusion}
\input{conclusion}

% ---------------------------------------------------------------------------
% ACKNOWLEDGEMENTS
% ---------------------------------------------------------------------------
% \section*{Acknowledgements}\label{sec:ack}
% \input{ack}

@@
